% !TEX root = ../main.tex
\begin{publications}{99}
\ifphd
\addcontentsline{toe}{chapter}{\bf Papers published in the period of doctor education}
\else
\addcontentsline{toe}{chapter}{\bf Papers published in the period of master education}
\fi
%\thispagestyle{empty}

%%盲审时候提交版本
\if 0
\item[1.]
**. ------------ [J]. \textit{IEEE Transactions on Multimedia (TMM)}.  (第一作者， Q1， 中科院一区， 影响因子：7.3，已发表)

\item[2.]
**. ------------ [J]. \textit{IEEE Transactions on Biomedical Engineering (T-BME)}.  (第一作者， Q1，中科院二区， 影响因子：4.6，已发表)

\item[3.]
**. ------------ [J]. \textit{IEEE Transactions on Image Processing (TIP)} (第一作者， Q1， 中科院一区， 影响因子：10.6，已录用)

\item[4.]
**. ------------ [J].\textit{npj Parkinsons Disease}. （第一作者， Q1， 中科院一区， 影响因子：8.7，审稿中）

\item[5.]
**. ------------ [J].\textit{图学学报}. (第一作者， 中文核心，已发表)

\item[6.]
**. ------------ [C]. \textit{ML4CS}.  (共一作者， EI会议， EI检索，已发表)


\item[7.]
**. ------------ [J]. \textit{放射学实践}. (第二作者，中文核心， 已发表)

\fi



%%正式答辩后提交版本
%\if 0
\item[1.]
\textbf{Jinyu Wen}, Asad Khan, Amei Chen, Weilong Peng, Meie Fang, Ping Li and C. L. Philip Chen. High-Quality Fusion and Visualization for MR-PET Brain Tumor Images via Multi-Dimensional Features[J]. \textit{IEEE Transactions on Image Processing}, May 2024, published online, DOI: 10.1109/TIP.2024.3404660. (中科院一区，IF:10.6)

\item[2.]
\textbf{Jinyu Wen}, Feiwei Qin, Jiao Du, Meie Fang, Ping Li, Xinhua Wei, and C. L. Philip Chen. MsgFusion: Medical Semantic Guided Two-Branch Network for Multi-Modal Brain Image Fusion[J]. \textit{IEEE Transactions on Multimedia}, 26, 944-957, 2024.  (中科院一区，IF:7.3)

\item[3.]
\textbf{Jinyu Wen}, Yang Li, Meie Fang, Lei Zhu, David Dagan Feng, and Ping Li. Fine-Grained and Multiple Classification for Alzheimer's Disease with Wavelet Convolution Unit Network[J]. \textit{IEEE Transactions on Biomedical Engineering}, 70(9), 2592-2603, 2023.  (中科院二区，IF:4.6)

\item[4.]
\textbf{Jinyu Wen}, Amei Chen, Meie Fang, Xinhua Wei. Computer aided prediction of Parkinson's disease progression by machine learning from cross-modal data[J]. \textit{npj Parkinsons Disease}. （审稿中，中科院一区，IF:8.7）

\item[5.]
\textbf{温金玉}, 方美娥. 基于机器学习的计算机辅助帕金森疾病分类预测研究综述[J].\textit{图学学报}, 44(6), 2023. (中文核心)

\item[6.]
Danyang Yao, \textbf{Jinyu Wen}, Amei Chen, Meie Fang, Xinhua Wei, Zhigeng Pan. Trimodal Fusion Network Combined Global-Local Feature Extraction Strategy and Spatial-Frequency Fusion Strategy. \textit{ML4CS} (3): 212-226, 2022.  (共一，EI会议收录)

\item[7.]
陈阿梅, \textbf{温金玉}, 黄俊祥, 成小芳, 阮秀杭, 方美娥, 魏新华. 血管周围间隙扩散张量成像指数在评估帕金森病患者脑类淋巴系统功能中的价值[J]. \textit{放射学实践}, 8: 954-959, 2022. (中文核心)

\item[8.]
Jipeng Hu, \textbf{Jinyu Wen}, Meie Fang. A survey on adversarial attack and defense of deep learning models for medical image recognition[J]. \textit{Metaverse} 2023; 4(1): 17，2023. 

\item[9.]
Asad Khan, Sakander Hayat, Muhammad Ahmad, \textbf{Jinyu Wen}, Muhammad Umar Farooq, Meie Fang, Wenchao Jiang. (2022). Cross‐modal retrieval based on deep regularized hashing constraints[J]. \textit{International Journal of Intelligent Systems}, 37(9), 6508-6530. (中科院 二区，IF:7)


\item[10.]
方美娥, 姚丹阳, \textbf{温金玉}. 一种基于深度学习的多模态医学图像融合方法. 2022-08-05, ZL 202210936228.6
%\fi




\end{publications}
